---
title: 'METHOD IV: ESCALATION'
author: "LABARTHE, S."
date: "2026-04-30"
output: pdf_document
---

# CODEBOOK

* `mid` - Militarized Interstate Dispute onset the next year
* `country1` - First state in the dyad
* `country2` - Second state in the dyad
* `dyad_id` - Dyad Identifiant 
* `year` - Calendar year
* `fatality` - Fatality (0-1)
* `hostlev` - Hostility level
* `mindur` - Minimum dispute duration (log days)
* `maxdur` - Maximum dispute duration (log days)
* `recip` - Reciprocation
* `max_fatality` - Maximum fatality level across dyad
* `min_fatality` - Minimum fatality level across dyad
* `max_hiact` - Maximum highest action level across dyad (0-21 scale)
* `min_hiact` - Minimum highest action level across dyad (0-21 scale)
* `max_hostlev` - Maximum hostility level across dyad
* `min_hostlev` - Minimum hostility level across dyad
* `max_milex` - Maximum military expenditure (thousands USD)
* `min_milex` - Minimum military expenditure (thousands USD)
* `max_milper` - Maximum military personnel (thousands)
* `min_milper` - Minimum military personnel (thousands)
* `max_irst` - Maximum industrial capacity (0-100 scale)
* `min_irst` - Minimum industrial capacity (0-100 scale)
* `max_pec` - Maximum primary energy consumption (coal-ton equivalents)
* `min_pec` - Minimum primary energy consumption (coal-ton equivalents)
* `max_tpop` - Maximum total population (thousands)
* `min_tpop` - Minimum total population (thousands)
* `max_upop` - Maximum urban population (thousands)
* `min_upop` - Minimum urban population (thousands)
* `max_cinc` - Maximum Composite Index of National Capability (0-1 scale)
* `min_cinc` - Minimum Composite Index of National Capability (0-1 scale)
* `wbgdp2011est1` - GDP estimate country1 (2011 USD, log)
* `wbgdp2011est2` - GDP estimate country2 (2011 USD, log)
* `max_wbgdp2011est` - Maximum GDP across dyad (log)
* `min_wbgdp2011est` - Minimum GDP across dyad (log)
* `ratio_gdp` - Ratio of smaller to larger GDP
* `max_wbpopest` - Maximum population across dyad (log)
* `min_wbpopest` - Minimum population across dyad (log)
* `ratio_pop` - Ratio of smaller to larger population
* `max_sdpest` - Maximum share of dyadic population (log)
* `min_sdpest` - Minimum share of dyadic population (log)
* `max_wbgdppc2011est` - Maximum GDP per capita (2011 USD, log)
* `min_wbgdppc2011est` - Minimum GDP per capita (2011 USD, log)
* `ratio_gdppc` - Ratio of smaller to larger GDP per capita
* `max_mrgdppc` - Maximum Maddison GDP per capita
* `min_mrgdppc` - Minimum Maddison GDP per capita
* `max_sd_mrgdppc` - Maximum standard deviation of GDP per capita
* `min_sd_mrgdppc` - Minimum standard deviation of GDP per capita
* `max_pwtrgdp` - Maximum Penn World Table GDP
* `min_pwtrgdp` - Minimum Penn World Table GDP
* `max_sd_pwtrgdp` - Maximum standard deviation of PWT GDP
* `min_sd_pwtrgdp` - Minimum standard deviation of PWT GDP
* `max_pwtpop` - Maximum PWT population
* `min_pwtpop` - Minimum PWT population
* `max_sd_pwtpop` - Maximum standard deviation of PWT population
* `min_sd_pwtpop` - Minimum standard deviation of PWT population
* `trade_total` - Total trade flow (USD)
* `trade_dep1` - Trade dependence country1 (flow1/GDP1)
* `trade_dep2` - Trade dependence country2 (flow2/GDP2)
* `max_trade_dep` - Maximum trade dependence across dyad
* `min_trade_dep` - Minimum trade dependence across dyad
* `ratio_trade` - Ratio of smaller to larger trade dependence
* `max_democracy` - Maximum democracy score across dyad (0-1)
* `min_democracy` - Minimum democracy score across dyad (0-1)
* `joint_democracy` - Both states democratic
* `land_contiguity` - Shared land border
* `water_contiguity` - Shared water border (< 24 miles)
* `capdist` - Capital distance (km)
* `capdist_log` - Log-transformed capital distance
* `atop_alliance` - Any formal alliance
* `atop_defense` - Defense pact
* `atop_neutral` - Neutrality agreement
* `atop_nonagg` - Non-aggression treaty
* `atop_consul` - Consultation/entente agreement
* `dyadigos` - Number of shared IGO memberships
* `max_leadertransition` - Maximum leadership transition
* `min_leadertransition` - Minimum leadership transition
* `max_irregular` - Maximum irregular leadership change
* `min_irregular` - Minimum irregular leadership change
* `max_n_leaders` - Maximum number of leaders
* `min_n_leaders` - Minimum number of leaders
* `positional` - Positional rivalry
* `spatial` - Spatial rivalry
* `ideological` - Ideological rivalry
* `interventionary` - Interventionary rivalry
* `ongoingrivalry` - Ongoing strategic rivalry
* `mid_count_5yr` - Number of MID onsets in previous 5 years
* `mid_lag1` - MID onset previous year
* `mid_lag2` - MID onset 2 years prior
* `mid_lag3` - MID onset 3 years prior
* `peace_years` - Consecutive years without MID onset
* `cowmaj1` - Country1 COW major power
* `cowmaj2` - Country2 COW major power
* `major1` - At least one major power in dyad
* `majors2` - Both states major powers

# DATA CONSTRUCTION
```{r, eval = F, message = F, warning = F}
rm(list=ls())
library(peacesciencer)
library(vdemdata)      
library(unvotes)
library(dplyr)
library(tidyr)

library(countrycode)
library(haven)    
library(readxl)

# Base
dyad_base <- create_dyadyears(
  system   = "cow",   
  directed = FALSE,     
  subset_years = c(1946:2012)  
) %>%
  mutate(
    country1 = countrycode(ccode1, "cown", "country.name"),
    country2 = countrycode(ccode2, "cown", "country.name"),
    dyad_id = paste0(pmin(country1, country2), "-", 
                       pmax(country1, country2))
  )

# MIDS
dyad_base <- dyad_base %>%
  add_cow_mids() %>% 
  mutate(
    across(everything(), ~replace(., is.na(.), 0)),
    fatality = ifelse(fatality == -9, 1, fatality), 
    fatality1 = ifelse(fatality1 == -9, 1, fatality1), 
    fatality2 = ifelse(fatality2 == -9, 1, fatality2),
    max_fatality = pmax(fatality1, fatality2),
    min_fatality = pmin(fatality1, fatality2),
    max_hiact = pmax(hiact1, hiact2),
    min_hiact = pmin(hiact1, hiact2),
    max_hostlev = pmax(hostlev1, hostlev2),
    min_hostlev = pmin(hostlev1, hostlev2), 
    mid = as.integer(!is.na(hostlev) & hostlev >= 1)
  ) %>% 
  select(-cowmidongoing, -dispnum, -cowmidonset, -sidea1, -sidea2, 
         -fatality1, -fatality2, -fatalpre1, -fatalpre2, 
         -hiact1, -hiact2, -hostlev1, -hostlev2, -orig1, -orig2, -stmon)

# CINC
dyad_base <- dyad_base %>%
  add_nmc() %>% 
  mutate(
    max_milex = pmax(milex1, milex2), 
    min_milex = pmin(milex1, milex2), 
    max_milper = pmax(milper1, milper2), 
    min_milper = pmin(milper1, milper2), 
    max_irst = pmax(irst1, irst2), 
    min_irst = pmin(irst1, irst2), 
    max_pec = pmax(pec1, pec2), 
    min_pec = pmin(pec1, pec2),   
    max_tpop = pmax(tpop1, tpop2), 
    min_tpop = pmin(tpop1, tpop2), 
    max_upop = pmax(upop1, upop2), 
    min_upop = pmin(upop1, upop2), 
    max_cinc = pmax(cinc1, cinc2), 
    min_cinc = pmin(cinc1, cinc2), 
    maxdur = log(maxdur + 1) , 
    mindur = log(mindur + 1) 
    ) %>% 
  select(-milex1, -milper1, -irst1, -pec1, -tpop1, -upop1, -cinc1, 
         -milex2, -milper2, -irst2, -pec2, -tpop2, -upop2, -cinc2)

# Economy
dyad_base <- dyad_base %>%
  add_sdp_gdp() %>% 
  mutate(
    max_wbgdp2011est = pmax(wbgdp2011est1, wbgdp2011est2), 
    min_wbgdp2011est = pmin(wbgdp2011est1, wbgdp2011est2), 
    
    ratio_gdp      = min_wbgdp2011est / max_wbgdp2011est,
    
    max_wbpopest = pmax(wbpopest1, wbpopest2), 
    min_wbpopest = pmin(wbpopest1, wbpopest2), 
    
    ratio_pop    = min_wbpopest / max_wbpopest,

    max_sdpest = pmax(sdpest1, sdpest2), 
    min_sdpest = pmin(sdpest1, sdpest2), 
    max_wbgdppc2011est = pmax(wbgdppc2011est1, wbgdppc2011est2), 
    min_wbgdppc2011est = pmin(wbgdppc2011est1, wbgdppc2011est2), 
    
    ratio_gdppc    = min_wbgdppc2011est / max_wbgdppc2011est
    ) %>% 
  select(-wbpopest1, -sdpest1, -wbgdppc2011est1,
         -wbpopest2, -sdpest2, -wbgdppc2011est2)

dyad_base <- dyad_base %>% 
  add_sim_gdp_pop() %>% 
  mutate(
    max_mrgdppc = pmax(mrgdppc1, mrgdppc2), 
    min_mrgdppc = pmin(mrgdppc1, mrgdppc2), 
    max_sd_mrgdppc = pmax(sd_mrgdppc1, sd_mrgdppc2), 
    min_sd_mrgdppc = pmin(sd_mrgdppc1, sd_mrgdppc2), 
    max_pwtrgdp = pmax(pwtrgdp1, pwtrgdp2), 
    min_pwtrgdp = pmin(pwtrgdp1, pwtrgdp2), 
    max_sd_pwtrgdp = pmax(sd_pwtrgdp1, sd_pwtrgdp2), 
    min_sd_pwtrgdp = pmin(sd_pwtrgdp1, sd_pwtrgdp2),   
    max_pwtpop = pmax(pwtpop1, pwtpop2), 
    min_pwtpop = pmin(pwtpop1, pwtpop2), 
    max_sd_pwtpop = pmax(sd_pwtpop1, sd_pwtpop2), 
    min_sd_pwtpop = pmin(sd_pwtpop1, sd_pwtpop2)
    ) %>% 
  select(-mrgdppc1, -sd_mrgdppc1, -pwtrgdp1, -sd_pwtrgdp1, -pwtpop1, -sd_pwtpop1, 
         -mrgdppc2, -sd_mrgdppc2, -pwtrgdp2, -sd_pwtrgdp2, -pwtpop2, -sd_pwtpop2)

# Trade
dyad_base <- dyad_base %>%
  add_cow_trade() %>%
  mutate(
    trade_total    = flow1 + flow2,
    trade_dep1     = flow1 / (wbgdp2011est1 + 1), 
    trade_dep2     = flow2 / (wbgdp2011est2 + 1),  
    max_trade_dep  = pmax(trade_dep1, trade_dep2),
    min_trade_dep  = pmin(trade_dep1, trade_dep2), 
    ratio_trade    = min_trade_dep / (max_trade_dep + 0.00001)
  ) %>% 
  select(-flow1, -flow2, -smoothflow1, -smoothflow2)

# Domestic politics
dyad_base <- dyad_base %>%
  add_democracy() %>% 
  mutate(
    max_democracy  = pmax(v2x_polyarchy1, v2x_polyarchy2),
    min_democracy  = pmin(v2x_polyarchy1, v2x_polyarchy2),
    joint_democracy = ifelse(max_democracy > 0.5 & min_democracy > 0.5, 1, 0)
  ) %>% 
  select(-v2x_polyarchy1,-v2x_polyarchy2, -euds1, -euds2, -aeuds1, -aeuds2, -polity21, -polity22)

dyad_base <- dyad_base %>%
  add_archigos() %>%
  mutate(
    max_leadertransition = pmax(leadertransition1, leadertransition2, na.rm = TRUE),
    min_leadertransition = pmin(leadertransition1, leadertransition2, na.rm = TRUE),
    max_irregular        = pmax(irregular1, irregular2, na.rm = TRUE),
    min_irregular        = pmin(irregular1, irregular2, na.rm = TRUE),
    max_n_leaders        = pmax(n_leaders1, n_leaders2, na.rm = TRUE),
    min_n_leaders        = pmin(n_leaders1, n_leaders2, na.rm = TRUE)) %>%
  select(-leadertransition1, -irregular1, -n_leaders1, 
         -leadertransition2, -irregular2, -n_leaders2) 

# Geography

dyad_base <- dyad_base %>%
  add_contiguity() %>%
  mutate(
    land_contiguity   = as.integer(!is.na(conttype) & conttype == 1), 
    water_contiguity  = as.integer(!is.na(conttype) & conttype == 2 | conttype ==3)
  ) %>% 
  select(-conttype)

dyad_base <- dyad_base %>%
  add_minimum_distance() %>%
  mutate(mindist_log = log(mindist + 1)) %>%
  select(-mindist)

dyad_base <- dyad_base %>%
  add_capital_distance() %>%
  mutate(
    capdist_log = log(capdist + 1)  
  )

# International 
dyad_base <- dyad_base %>% 
  add_atop_alliance() %>% 
  mutate(atop_alliance = ifelse(is.na(atop_alliance), 0, atop_alliance)) %>% 
  select(-atop_offense)

dyad_base <- dyad_base %>%
  add_igos() 

dyad_base <- dyad_base %>%
  add_strategic_rivalries() %>% 
  select(-jan1obsid1, -dec31obsid1, -jan1obsid2, -dec31obsid2, -tssr_id, )

# Final adjustments
dyad_base <- dyad_base %>%
  arrange(country1, country2, year) %>%
  group_by(ccode1, ccode2) %>%
  mutate(
    mid_count_5yr   = zoo::rollsum(mid, k = 5, 
                                   fill = NA, 
                                   align = "right"),

    mid_lag1        = lag(mid, 1),
    mid_lag2        = lag(mid, 2),
    mid_lag5        = lag(mid, 5),

    peace_years     = sequence(rle(mid == 0)$lengths) * (mid == 0)
  ) %>%
  ungroup()

attr(dyad_base, "ps_data_type") <- "dyad_year"
attr(dyad_base, "ps_system")    <- "cow"

dyad_base <- dyad_base %>% 
  add_cow_majors() %>%
  mutate(
    major1  = as.integer(cowmaj1 == 1 | cowmaj2 == 1),  
    majors2 = as.integer(cowmaj1 == 1 & cowmaj2 == 1)   
  )

# Lag + 1
keep_cols <- c("country1", "country2", "year", "dyad_id", "mid")

dyad_base <- dyad_base %>%
  group_by(country1, country2) %>%       
  mutate(
    across(
      -any_of(keep_cols),               
      ~ lag(.x, n = 1)
    )
  ) %>%
  ungroup()

data <- dyad_base %>% 
  filter(year >= 1992 & year <= 2012) %>% 
  mutate(mid = as.factor(mid)) %>% 
  select(mid, country1, country2, dyad_id, year, everything(), -ccode1, -ccode2) %>% 
  arrange(country1, country2, year) %>% 
  na.omit()

test <- dyad_base %>% 
  filter(year >= 1992 & year <= 2012) %>% 
  mutate(mid = as.factor(lead(mid, 1))) %>% 
  select(mid, country1, country2, dyad_id, year, everything(), -ccode1, -ccode2) %>% 
  arrange(country1, country2, year)

length(unique(dyad_base$dyad_id))

length(unique(data$dyad_id))

saveRDS(data, "data.rds")
```

# NEXT
```{r}
dyad_base <- dyad_base %>%
  add_vdem(vars = c(
    "v2lgfemleg",       # % women in lower chamber of parliament
    "v2x_corr",         # political corruption index
    "v2x_civlib",       # civil liberties index
    "v2x_freexp_altinf" # freedom of expression & alternative info
  )) %>%
  mutate(
    max_femleg  = pmax(v2lgfemleg1, v2lgfemleg2, na.rm = TRUE),
    min_femleg  = pmin(v2lgfemleg1, v2lgfemleg2, na.rm = TRUE),
    max_corr    = pmax(v2x_corr1, v2x_corr2, na.rm = TRUE),
    min_corr    = pmin(v2x_corr1, v2x_corr2, na.rm = TRUE),
    max_civlib  = pmax(v2x_civlib1, v2x_civlib2, na.rm = TRUE),
    min_civlib  = pmin(v2x_civlib1, v2x_civlib2, na.rm = TRUE)
  ) %>%
  select(-v2lgfemleg1, -v2lgfemleg2, -v2x_corr1, -v2x_corr2,
         -v2x_civlib1, -v2x_civlib2,
         -v2x_freexp_altinf1, -v2x_freexp_altinf2)

# V-Party: governing party ideology & nationalism
dyad_base <- dyad_base %>%
  add_vparty() %>%
  mutate(
    # Left-right ideology of governing party
    max_govideology   = pmax(v2pariglef1, v2pariglef2, na.rm = TRUE),
    min_govideology   = pmin(v2pariglef1, v2pariglef2, na.rm = TRUE),
    ideological_dist  = abs(v2pariglef1 - v2pariglef2),  # ideological distance between govts

    # Nationalism of governing party
    max_nationalism   = pmax(v2panas1, v2panas2, na.rm = TRUE),
    min_nationalism   = pmin(v2panas1, v2panas2, na.rm = TRUE),
    nationalism_dist  = abs(v2panas1 - v2panas2)  
  ) %>%
  select(-v2pariglef1, -v2pariglef2, -v2panas1, -v2panas2)
```

