---
title: "METHOD IV PROJECT"
author: "LABARTHE, S."
date: '2026'
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

# 1. DATA
```{r data, message=FALSE, warning=FALSE}
library(dplyr)
library(readr)
library(rsample)
library(tidymodels)
library(glmnet)
library(ranger)
library(xgboost)
library(ggplot2)
library(tidyr)
library(doParallel)

set.seed(42)
tidymodels_prefer()

data <- read_rds("https://raw.githubusercontent.com/SimonLabarthe/Militarised-Interstate-Dispute-Predicted-Probabilities-Dataset/main/data.rds") %>%
  rename(y = mid)

data_split <- initial_split(data, prop = 0.9, strata = y)
train <- training(data_split)
test  <- testing(data_split)

cat("Train:", nrow(train), "| Test:", nrow(test), "\n")
cat("Class balance (train):\n")
print(prop.table(table(train$y)) %>% round(3))

# Models
fit.logit <- readRDS("models/fit.logit.rds")
fit.elnet <- readRDS("models/fit.elnet.rds")
fit.rf <- readRDS("models/fit.rf.rds")
fit.xgb <- readRDS("models/fit.xgb.rds")

# Predictions
preds <- readRDS("models/preds.rds")
results <- readRDS("models/results.rds")

tune_rf <- readRDS("models/tune_rf.rds")
```

---

# 2. GLM FIT (no tuning needed)
```{r fit_logit, message=FALSE, warning=FALSE}
rec_base <- recipe(y ~ ., data = train) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors())

spec_logit <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

wf_logit <- workflow() %>% 
  add_recipe(rec_base)  %>% 
  add_model(spec_logit)

fit.logit <- wf_logit %>% 
  fit(data = train)

saveRDS(fit.logit, "models/fit.logit.rds")
fit.logit
```

---

# 3. HYPERPARAMETER TUNING
## a. Elastic Net
```{r tune_elnet, message=FALSE, warning=FALSE}
# ID continuous variables
continuous_cols <- train %>%
  select(where(is.numeric)) %>%
  select(where(~ length(unique(na.omit(.x))) > 2)) %>%
  names()

length(continuous_cols)
print(continuous_cols)

rec_elnet <- recipe(y ~ ., data = train) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_poly(all_of(continuous_cols), degree = 2, keep_original_cols = FALSE) %>%
  step_zv(all_predictors())

spec_elnet <- logistic_reg(
  penalty = tune(),   # L1+L2 regularisation strength
  mixture = tune()    # 0 = ridge, 1 = lasso, 0.5 = elastic net
) %>%
  set_engine("glmnet") %>%
  set_mode("classification")

wf_elnet  <- workflow() %>% 
  add_recipe(rec_elnet) %>% 
  add_model(spec_elnet)

# 5-fold CV, stratified on y
cv_folds <- vfold_cv(train, v = 5, strata = y)

# Parallelise all tuning
cl <- makeCluster(parallel::detectCores() - 1)
registerDoParallel(cl)

grid_elnet <- grid_regular(
  penalty(range = c(-4, 0)),   # 10^-4 to 10^0
  mixture(range = c(0, 1)),
  levels = 6
)

tune_elnet <- tune_grid(
  wf_elnet,
  resamples = cv_folds,
  grid      = grid_elnet,
  metrics   = metric_set(roc_auc, accuracy, f_meas)
)

autoplot(tune_elnet) + 
  labs(title = "Elastic Net — Hyperparameter Tuning") +
  theme_bw()
  
best_elnet    <- select_best(tune_elnet, metric = "roc_auc")
fit.elnet     <- finalize_workflow(wf_elnet, best_elnet) %>% fit(data = train)
saveRDS(fit.elnet, "models/fit.elnet.rds")
cat("Best elastic net — penalty:", best_elnet$penalty, 
    " | mixture:", best_elnet$mixture, "\n")

stopCluster(cl)
registerDoSEQ()
```

## b. Random Forest — ROC AUC tuning
```{r tune_rf, message=FALSE, warning=FALSE}
rec_base <- recipe(y ~ ., data = train) %>%
  step_zv(all_predictors()) %>% 
  step_normalize(all_numeric_predictors())

spec_rf <- rand_forest(
  mtry  = tune(),     
  trees = 100,
  min_n = tune()     
) %>%
  set_engine("ranger", 
             importance = "impurity", 
             seed = 42, 
             replace = FALSE,
             sample_fraction = 0.5) %>%
  set_mode("classification")

wf_rf <- workflow() %>% 
  add_recipe(rec_base)  %>% 
  add_model(spec_rf)

# 5-fold CV, stratified on y
cv_folds <- bootstraps(train, times = 15, strata = y)

# Parallelise all tuning
cl <- makeCluster(parallel::detectCores() - 1)
registerDoParallel(cl)

grid_rf <- grid_random(
  mtry(range = c(3, 20)),
  min_n(range = c(5, 50)),
  size = 15
)

tune_rf <- tune_grid(
  wf_rf,
  resamples = cv_folds,
  grid      = grid_rf,
  metrics   = metric_set(roc_auc, accuracy, f_meas)
)
saveRDS("models/tune_rf.rds")

autoplot(tune_rf) + 
  labs(title = "Random Forest — Hyperparameter Tuning (ROC AUC)") +
  theme_minimal()

best_rf   <- select_best(tune_rf, metric = "roc_auc")
fit.rf    <- finalize_workflow(wf_rf, best_rf) %>% fit(data = train)
saveRDS(fit.rf, "models/fit.rf.rds")
cat("Best RF — mtry:", best_rf$mtry, " | min_n:", best_rf$min_n, "\n")

stopCluster(cl)
registerDoSEQ()
```

## c. XGBoost
```{r tune_xgb, message=FALSE, warning=FALSE}
rec_base <- recipe(y ~ ., data = train) %>%
  step_zv(all_predictors()) %>%                      
  step_normalize(all_numeric_predictors())

spec_xgb <- boost_tree(
  trees      = 100,
  tree_depth = tune(),
  learn_rate = tune(),
  min_n      = tune()
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

wf_xgb    <- workflow() %>% 
  add_recipe(rec_base)  %>% 
  add_model(spec_xgb)

grid_xgb <- grid_random(
  tree_depth(range = c(3, 8)),
  learn_rate(range = c(-3, -1)),  
  min_n(range      = c(5, 50)),
  size = 15
)

# 5-fold CV, stratified on y
cv_folds <- bootstraps(train, times = 15, strata = y)

# Parallelise all tuning
cl <- makeCluster(parallel::detectCores() - 1)
registerDoParallel(cl)

tune_xgb <- tune_grid(
  wf_xgb,
  resamples = cv_folds,
  grid      = grid_xgb,
  metrics   = metric_set(roc_auc, 
                         accuracy, 
                         f_meas)
)

autoplot(tune_xgb) + 
  labs(title = "XGBoost — Hyperparameter Tuning") +
  theme_minimal()

best_xgb  <- select_best(tune_xgb, metric = "roc_auc")
fit.xgb   <- finalize_workflow(wf_xgb, best_xgb) %>% fit(data = train)
saveRDS(fit.xgb, "models/fit.xgb.rds")
cat("Best XGBoost — depth:", best_xgb$tree_depth, 
    " | lr:", round(best_xgb$learn_rate, 4),
    " | min_n:", best_xgb$min_n, "\n")

stopCluster(cl)
registerDoSEQ()
```

## d. Support Vector Machine (SVM) !! TOO LONG TO COMPUTE
```{r tune_svm, message=FALSE, warning=FALSE}
library(kernlab)  # for SVM
library(kknn)     # for KNN

# Recipe for SVM (normalization is critical for SVM)
rec_svm <- recipe(y ~ ., data = train) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors())

# SVM specification with RBF kernel
spec_svm <- svm_rbf(
  cost      = tune(),      # regularization parameter
  rbf_sigma = tune()       # kernel parameter (gamma)
) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

wf_svm <- workflow() %>% 
  add_recipe(rec_svm) %>% 
  add_model(spec_svm)

# Tuning grid for SVM
grid_svm <- grid_regular(
  cost(range = c(-2, 2)),      # 0.01 to 100
  rbf_sigma(range = c(-3, 1)), # 0.001 to 10
  levels = 6
)

# 5-fold CV stratified on y
cv_folds_svm <- vfold_cv(train, v = 2, strata = y)

# Setup parallelization
cl <- makeCluster(parallel::detectCores() - 1)
registerDoParallel(cl)

tune_svm <- tune_grid(
  wf_svm,
  resamples = cv_folds_svm,
  grid      = grid_svm,
  metrics   = metric_set(roc_auc, accuracy, f_meas),
  control   = control_grid(verbose = FALSE)
)

autoplot(tune_svm) + 
  labs(title = "SVM (RBF Kernel) — Hyperparameter Tuning") +
  theme_bw()

best_svm  <- select_best(tune_svm, metric = "roc_auc")
fit.svm   <- finalize_workflow(wf_svm, best_svm) %>% fit(data = train)
saveRDS(fit.svm, "models/fit.svm.rds")
cat("Best SVM — cost:", round(best_svm$cost, 3), 
    " | rbf_sigma:", round(best_svm$rbf_sigma, 4), "\n")

stopCluster(cl)
registerDoSEQ()
```

## e. K-Nearest Neighbors (KNN)
```{r}

# Recipe for KNN (normalization is critical due to distance-based algorithm)
rec_knn <- recipe(y ~ ., data = train) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors())

# KNN specification
spec_knn <- nearest_neighbor(
  neighbors = tune(),      # k: number of neighbors
  weight_func = tune()     # weighting function (e.g., "rectangular", "triangular", "epanechnikov")
) %>%
  set_engine("kknn") %>%
  set_mode("classification")

wf_knn <- workflow() %>% 
  add_recipe(rec_knn) %>% 
  add_model(spec_knn)

# Tuning grid for KNN
grid_knn <- grid_regular(
  neighbors(range = c(3, 51)),     # odd numbers preferred to avoid ties
  weight_func(),
  levels = c(neighbors = 12, weight_func = 5)
)

# 5-fold CV stratified on y
cv_folds_knn <- vfold_cv(train, v = 2, strata = y)

# Parallelization
cl <- makeCluster(parallel::detectCores() - 1)
registerDoParallel(cl)

tune_knn <- tune_grid(
  wf_knn,
  resamples = cv_folds_knn,
  grid      = grid_knn,
  metrics   = metric_set(roc_auc, accuracy, f_meas),
  control   = control_grid(verbose = FALSE)
)

autoplot(tune_knn) + 
  labs(title = "KNN — Hyperparameter Tuning") +
  theme_bw()

best_knn   <- select_best(tune_knn, metric = "roc_auc")
fit.knn    <- finalize_workflow(wf_knn, best_knn) %>% fit(data = train)
saveRDS(fit.knn, "models/fit.knn.rds")
cat("Best KNN — neighbors:", best_knn$neighbors, 
    " | weight_func:", best_knn$weight_func, "\n")

stopCluster(cl)
registerDoSEQ()
```

---

# 4. TEST SET PREDICTIONS

```{r predictions_updated}
# Helper functions
get_class <- function(fit) predict(fit, new_data = test, type = "class")$.pred_class
get_prob  <- function(fit) predict(fit, new_data = test, type = "prob")$.pred_1

preds <- tibble(
  truth      = test$y,
  # classes
  logit_cls  = get_class(fit.logit),
  elnet_cls  = get_class(fit.elnet),
  rf_cls     = get_class(fit.rf),
  xgb_cls    = get_class(fit.xgb),
  #svm_cls    = get_class(fit.svm),
  #knn_cls    = get_class(fit.knn),
  # probabilities of class "1" (MID = 1)
  logit_prob = get_prob(fit.logit),
  elnet_prob = get_prob(fit.elnet),
  rf_prob    = get_prob(fit.rf),
  xgb_prob   = get_prob(fit.xgb),
  #svm_prob   = get_prob(fit.svm),
  #knn_prob   = get_prob(fit.knn)
)

saveRDS(preds, "models/preds.rds")
```

---

# 5. EVALUATION

## a. Confusion Matrices

```{r conf_matrices}
models <- c("logit", "elnet", "rf", "xgb")
labels <- c("Logistic", "Elastic Net", "Random Forest", "XGBoost")

for (i in seq_along(models)) {
  cat("\n───", labels[i], "───\n")
  cls_col <- paste0(models[i], "_cls")
  print(conf_mat(preds, truth = truth, estimate = !!sym(cls_col))$table)
}
```

```{r conf_matrices_figure, fig.width=12, fig.height=10}
library(ggplot2)
library(paletteer)
plot_confusion_matrix_plasma_pct <- function(model_name, model_label, preds_df) {
  cls_col <- paste0(model_name, "_cls")
  
  cm_table <- conf_mat(preds_df, truth = truth, estimate = !!sym(cls_col))$table
  cm_df <- as.data.frame(cm_table)
  colnames(cm_df) <- c("Truth", "Prediction", "Freq")
  
  cm_df$Truth <- as.character(cm_df$Truth)
  cm_df$Prediction <- as.character(cm_df$Prediction)
  
  # Calculate row percentages
  cm_df <- cm_df %>%
    group_by(Truth) %>%
    mutate(Percentage = Freq / sum(Freq) * 100,
           Label = paste0(Freq, "\n(", round(Percentage, 1), "%)"))
  
  ggplot(cm_df, aes(x = Prediction, y = Truth, fill = Percentage)) +
    geom_tile(color = "white", linewidth = 1) +
    geom_text(aes(label = Label), size = 5, fontface = "bold", color = "white") +
    scale_fill_paletteer_c("grDevices::Inferno", 
                           name = "Row\nPercentage (%)",
                           limits = c(0, 100)) +
    scale_x_discrete(position = "top") +
    labs(title = model_label,
         x = "Truth",
         y = "Predicted MID") +
    theme_bw(base_size = 12) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      axis.text = element_text(size = 11, face = "bold"),
      axis.title = element_text(size = 11),
      panel.grid = element_blank(),
      legend.position = "right",
      legend.key.width = unit(0.5, "cm"),
      legend.key.height = unit(1.5, "cm")
    ) +
    coord_fixed()
}

# Create plots with percentage fill
p1_pct <- plot_confusion_matrix_plasma_pct("logit", "Logistic", preds)
p2_pct <- plot_confusion_matrix_plasma_pct("elnet", "Elastic Net", preds)
p3_pct <- plot_confusion_matrix_plasma_pct("rf", "Random Forest", preds)
p4_pct <- plot_confusion_matrix_plasma_pct("xgb", "XGBoost", preds)

# Combine
library(patchwork)
combined_plot_pct <- (wrap_elements(p1_pct) + wrap_elements(p2_pct)) /
                     (wrap_elements(p3_pct) + wrap_elements(p4_pct)) +
  plot_annotation(
    theme = theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
      plot.subtitle = element_text(hjust = 0.5, size = 10, color = "gray40")
    )
  ) &
  theme(legend.position = "right")

# Save
pdf("graphs/confusion_matrices_plasma_pct.pdf", width = 12, height = 6)
print(combined_plot_pct)
dev.off()

combined_plot_pct
```


## b. Accuracy, Precision, Recall per model

```{r metrics_table}
get_metrics <- function(cls_col, prob_col, model_name) {
  tibble(
    Model     = model_name,
    Accuracy  = accuracy(preds,  truth, estimate = !!sym(cls_col))$.estimate,
    Precision = precision(preds, truth, estimate = !!sym(cls_col),
                          event_level = "second")$.estimate,
    Recall    = recall(preds,    truth, estimate = !!sym(cls_col),
                       event_level = "second")$.estimate,
    F1        = f_meas(preds,    truth, estimate = !!sym(cls_col),
                       event_level = "second")$.estimate,
    AUC       = roc_auc(preds,   truth, !!sym(prob_col),
                        event_level = "second")$.estimate
  )
}

results <- bind_rows(
  get_metrics("logit_cls", "logit_prob", "Logistic"),
  get_metrics("elnet_cls", "elnet_prob", "Elastic Net"),
  get_metrics("rf_cls",    "rf_prob",    "Random Forest"),
  get_metrics("xgb_cls",   "xgb_prob",  "XGBoost")
) %>%
  mutate(across(where(is.numeric), ~round(., 3)))

saveRDS(results, "models/results.rds")

library(stargazer)
stargazer(results, 
          summary = FALSE,
          rownames = FALSE,
          title = "Model Performance Comparison",
          digits = 3,
          type = "latex")
```

## c. ROC Curves — all models

```{r roc_curves, fig.width=8, fig.height=6}
# Build tidy ROC data for all four models
roc_data <- bind_rows(
  roc_curve(preds, truth, logit_prob, event_level = "second") %>% mutate(model = "Logistic"),
  roc_curve(preds, truth, elnet_prob, event_level = "second") %>% mutate(model = "Elastic Net"),
  roc_curve(preds, truth, rf_prob,    event_level = "second") %>% mutate(model = "Random Forest"),
  roc_curve(preds, truth, xgb_prob,   event_level = "second") %>% mutate(model = "XGBoost")
)

# Add AUC to legend labels
auc_labels <- results %>%
  mutate(label = paste0(Model, "  (AUC = ", round(AUC, 3), ")")) %>%
  select(model = Model, label)

model_colors <- paletteer_c("grDevices::Inferno", n = 4)
names(model_colors) <- levels(results_long$Model)

roc_data <- roc_data %>% left_join(auc_labels, by = "model")

pdf("graphs/roc_curves.pdf", width = 10, height = 6)
ggplot(roc_data, aes(x = 1 - specificity, y = sensitivity, colour = model)) +
  geom_line(linewidth = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey60") +
  scale_colour_manual(
    values = model_colors,
    labels = setNames(auc_labels$label, auc_labels$model),
    name = NULL
  ) +
  labs(
    x       = "False Positive Rate (1 − Specificity)",
    y       = "True Positive Rate (Sensitivity)"
  ) +
  theme_bw() +
  theme(
    legend.position = c(0.75, 0.25),
    legend.background = element_rect(fill = "white", linewidth = 0.2),
    legend.text = element_text(size = 9),
    legend.key.size = unit(0.8, "cm")
  )

dev.off()
```

## d. Performance Barplots — Accuracy, Precision, Recall
```{r barplots, fig.width=12, fig.height=5}
library(paletteer)

results_long <- results %>%
  pivot_longer(-Model, names_to = "Metric", values_to = "Value") %>%
  filter(Metric %in% c("Accuracy", "Precision", "Recall", "AUC")) %>%
  mutate(Metric = factor(Metric, levels = c("Accuracy", "Precision", "Recall", "AUC")),
         Model  = factor(Model,  levels = c("Logistic", "Elastic Net",
                                            "Random Forest", "XGBoost")))

# Generate plasma colors for models
model_colors <- paletteer_c("grDevices::Inferno", n = 4)
names(model_colors) <- levels(results_long$Model)

pdf("graphs/comparison_plasma.pdf", width = 10, height = 6)
comparison_short <- ggplot(results_long, aes(x = Model, y = Value, fill = Model)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = round(Value, 3)),
            vjust = -0.5, size = 3.8, fontface = "bold") +
  facet_wrap(~ Metric) +
  scale_fill_manual(values = model_colors) +
  scale_y_continuous(limits = c(0, 1.08)) +
  labs(
    x     = NULL,
    y     = NULL
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(face = "bold", size = 11),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10, face = "bold")
  )
dev.off()

comparison_short
```

---

# 6. TUNING RANDOM FOREST
```{r}
library(paletteer)

# Extract tuning results
rf_tune_results <- tune_rf %>% 
  collect_metrics() %>%
  filter(.metric == "roc_auc")

# Create a continuous palette for min_n values
# First, get unique min_n values sorted
min_n_values <- sort(unique(rf_tune_results$min_n))

# Generate colors from viridis plasma
plasma_colors <- paletteer_c("viridis::plasma", n = length(min_n_values))
names(plasma_colors) <- min_n_values

ggplot(rf_tune_results, 
       aes(x = mtry, 
           y = mean, 
           color = factor(min_n),
           group = min_n)) +
  geom_point(size = 2.5, alpha = 0.8) +
  geom_line(linewidth = 0.9) +
  scale_color_manual(values = plasma_colors, name = "min_n") +
  labs(
    title = "Random Forest — Hyperparameter Tuning",
    subtitle = "ROC AUC by mtry and min_n (Viridis Plasma)",
    x = "mtry (Number of Variables Tried at Each Split)",
    y = "ROC AUC"
  ) +
  scale_y_continuous(limits = c(0.70, 0.73), 
                     breaks = seq(0.70, 0.73, by = 0.002)) +
  theme_bw(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11, color = "gray40"),
    legend.position = "right",
    panel.grid.minor = element_blank(),
    legend.title = element_text(face = "bold")
  )
```


# 7. VARIABLE IMPORTANCE (RF)

```{r var_importance, fig.width=9, fig.height=6}
library(paletteer)
library(ggplot2)
library(dplyr)
library(labelled)

rf_engine <- fit.rf %>% 
  extract_fit_engine()

p <- 20

# Direct extraction of labels from your data object
importance_df <- tibble(
  variable   = names(rf_engine$variable.importance),
  importance = rf_engine$variable.importance
) %>%
  slice_max(importance, n = p) %>%
  # Get labels directly using labelled package
  mutate(variable_label = sapply(variable, function(x) {
    lbl <- attr(data[[x]], "label")
    if(is.null(lbl)) x else lbl
  }))

plasma_colors <- paletteer_c("grDevices::Inferno", n = p)

pdf("graphs/importance.pdf", width = 12, height = 6)  
importance <- ggplot(importance_df, aes(x = importance, 
                           y = reorder(variable_label, importance),
                           fill = importance)) +
  geom_col(width = 0.7) +
  scale_fill_gradientn(colors = plasma_colors, name = "Importance") +
  labs(x = "Mean Decrease in Gini Impurity",
       y = NULL) +
  theme_bw(base_size = 12) +
  theme(legend.position = "right",
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        axis.text.y = element_text(size = 10, hjust = 1))
print(importance)
dev.off()

importance
```

# 8. DATASET
```{r}
library(countrycode)
library(peacesciencer)
library(dplyr)
library(tidyr)

dyad_base <- create_dyadyears(
  system   = "cow",   
  directed = FALSE,     
  subset_years = c(1992:2012)  
) %>%
  mutate(
    country_i = countrycode(ccode1, "cown", "country.name"),
    country_j = countrycode(ccode2, "cown", "country.name"),
    dyad_id = paste0(pmin(country_i, country_j), "-", 
                       pmax(country_i, country_j))
  ) %>%
  mutate(
    country_i = as.character(country_i),
    country_j = as.character(country_j)
  )

mona_base <- create_stateyears(
  system   = "cow",   
  subset_years = c(1992:2012)  
) %>%
  mutate(
    country = countrycode(ccode, "cown", "country.name"),
    country = as.character(country)
  )

train_preds <- predict(fit.rf, new_data = train, type = "prob")$.pred_1
test_preds <- preds$rf_prob  

all_predictions <- bind_rows(
  train %>% select(y, country1, country2, dyad_id, year) %>% mutate(mid_p_hat = train_preds),
  test %>% select(y, country1, country2, dyad_id, year) %>% mutate(mid_p_hat = test_preds)
)

all_data_with_preds <- data %>%
  mutate(mid_p_hat = predict(fit.rf, new_data = data, type = "prob")$.pred_1) %>%
  select(dyad_id, country1, country2, year, mid_p_hat)

dyad_data <- dyad_base %>%
  left_join(
    all_data_with_preds %>% 
      mutate(country1 = as.character(country1),
             country2 = as.character(country2)),
    by = c("country_i" = "country1", 
           "country_j" = "country2", 
           "year" = "year")
  ) %>% 
  select(country_i, 
         country_j, 
         dyad_id = dyad_id.x, 
         year, 
         mid_p_hat) %>% 
  arrange(country_i, country_j, year)

monad_data <- dyad_data %>%
  filter(!is.na(mid_p_hat)) %>%
  pivot_longer(cols = c(country_i, country_j),
               values_to = "country") %>%
  group_by(country, year) %>%
  summarise(
    mid_p_hat = 1 - prod(1 - mid_p_hat),
    n_dyads = n(),
    .groups = "drop"
  ) %>%
  right_join(mona_base, by = c("country", "year")) %>%
  arrange(country, year) %>% 
  select(country, 
         year, 
         mid_p_hat)

saveRDS(dyad_data, "dyad_data.rds")
saveRDS(monad_data, "monad_data.rds")
```







